---
title: "Ocotillo Morphology Summary Statistics"
author: "Regressions and Kruskal-Wallis by Anastasia Bernat | PCA by Acacia Tsz So Tang"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
rm(list=ls())

library(ggplot2)
library(olsrr)
library(lme4)

library(tidyverse)
library(ggpubr)
library(rstatix)

library(FactoMineR)
library(factoextra)
library(readr)
library(corrplot)

dir = "/Users/anastasiabernat/Desktop/git_repositories/ocotillo-research/"
setwd(dir)

knitr::opts_chunk$set(echo = TRUE)
```


# Multivariate Modeling of Ocotillo Morphology

Analyses below are multivariate models of Fouquieria splendens morphology for ocotillos located on both a bajada and a plain in Organ Pipe National Monument, Arizona. All models were grouped by their response variable and ordered by their ascending AIC values. Dataset "ocos" indicates all individuals measured on the bajada and plain while "ocos_baj" indicates only the individuals measured on the bajada. Ocotillos located on the bajada were encoded with site = 0 while ocotillos on the plain were encoded with site = 1. Interspecific neighbor group is split between two types – shrub and cactus – where cactus = 0 and shrub = 1. 

## Reading and Cleaning the Data

```{r}
source("src/cleaning_data.R")
source("src/regression_output.R")
output_col = FALSE

all_data = clean_data("data/General_Oco_Data.csv", "data/Terminal_5Segs.csv")

ocos = all_data[[2]]
ocos_baj = ocos[ocos$Plant[1:20],]
```

## Functions

```{r}
### Diagnostic Plot Function

plot_diagnostic = function(m, d) {
  # m:  model generated by lm() or glm()
  # d:  data as a data.frame object
  
  d['fitted_values'] = m$fitted.values
  d['fitted_residuals'] = m$residuals
  
  plot = ggplot(data = d) + geom_point(aes(x = fitted_values, y = fitted_residuals)) +
  geom_hline(yintercept = 0) +
  labs(title = "Plot of Fitted Residuals against Fitted Values",
       x = "Fitted Values", y = "Fitted Residuals") 
  
  print(ols_test_normality(m))
  
  return(plot)
}
```


```{r}
### Renaming Regression Formulas Function

rename_regformula = function (eqs, response_var, predA, predB, predC, predD) {
  # eqs:              list of linear regression formals as characters
  # response_var:     response variable as a character
  # predA:            predictor variable A as a character
  # predB:            predictor variable B as a character
  # predC:            predictor variable C as a character
  # predD:            predictor variable D as a character
  for (i in 1:length(eqs)) {
  if (grepl(" C", eqs[i], fixed = TRUE)) {
    eqs[i] =  gsub("C", predC, eqs[i])
  }
  if (grepl(" A", eqs[i], fixed = TRUE)) {
    eqs[i] =  gsub("A", predA, eqs[i])
  }
  if (grepl(" B", eqs[i], fixed = TRUE)) {
    eqs[i] =  gsub("B", predB, eqs[i])
  }
  if (grepl(" D", eqs[i], fixed = TRUE)) {
    eqs[i] =  gsub("D", predD, eqs[i])
  }
  if (grepl( "R", eqs[i], fixed = TRUE)) {
    eqs[i] = gsub("R", response_var, eqs[i])
  }
  }
  return(eqs)
}
```


## Height 

### All Ocos

**Testing Covariates**

Only significant covariates are listed below:

```{r echo=FALSE}
m1 = lm(logHeight ~ Elevation_c, data=ocos)
m2 = lm(Height_c ~ Site, data=ocos)
m3 = lm(logHeight ~ Site, data=ocos)
m4 = lm(Height_c ~ TSegIQR_c, data=ocos)
m5 = lm(logHeight ~ TSegIQR_c, data=ocos)

tidy_regression(m1, output_col) 
tidy_regression(m2, output_col)
tidy_regression(m3, output_col)
tidy_regression(m4, output_col) 
tidy_regression(m5, output_col)
```

**Make Diagnostic Plots**

When the response variable is log transformed, residuals are independent.

```{r}
plot_diagnostic(m1, ocos)
#plot_diagnostic(m2, ocos) # categorical covariate
#plot_diagnostic(m3, ocos) # categorical covariate
plot_diagnostic(m4, ocos)
plot_diagnostic(m5, ocos)
```

**Model Comparisons**

**site, Elevation_c, TSeg_c**

```{r}
data<-data.frame(
  R=ocos$logHeight,
  A=ocos$Site,
  B=ocos$Elevation_c, 
  C=ocos$TSegIQR_c) 

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian lm 3-FF.R")
```

```{r}
cor(ocos$Site, ocos$Elevation_c)
```

Elevation and site were highly correlated (R = -0.85), which led to multicollinearity. In turn, we removed any interactions between elevation and site in the model comparison process in order to minimize relationships that were spurious. This eliminated m11 and m15 as the top models.

```{r}
anova(m7, m13, test='Chisq') # adding B*C does not improve fit
anova(m5, m7, test='Chisq') # Adding B does not improve fit
anova(m4, m7, test='Chisq') # Adding C does improve fit
anova(m1, m4, test='Chisq') # Adding B does not improve fit
anova(m3, m5, test='Chisq') # Adding C does improve fit
```

**Plot diagnostics**

All passed. Residuals are independent.

```{r}
plot_diagnostic(m5, ocos)
plot_diagnostic(m7, ocos)
plot_diagnostic(m4, ocos)
# plot_diagnostic(m1, ocos) # categorical covariate
plot_diagnostic(m3, ocos)
```

**Calculate likelihood and weights**

```{r}
# Calculate relative likelihood of each model which is exp( -0.5 * deltaAIC)

delta1 = 0 # abs(AIC(m5) - AIC(m5))
delta2 = abs(AIC(m5) - AIC(m7))
delta3 = abs(AIC(m5) - AIC(m4))
delta4 = abs(AIC(m5) - AIC(m1))
delta5 = abs(AIC(m5) - AIC(m3))

me1 <- exp( -0.5 * delta1)
me2 <- exp( -0.5 * delta2)
me3 <- exp( -0.5 * delta3)
me4 <- exp( -0.5 * delta4)
me5 <- exp( -0.5 * delta5)

# Calculate the Akaike weight for a model which is this value divided by the sum of these values across all models.

sum <- sum(c(me1, me2, me3, me4, me5))

weight1 = me1 / (sum) 
weight2 = me2 / (sum)
weight3 = me3 / (sum)
weight4 = me4 / (sum)
weight5 = me5 / (sum)
```

```{r}
summary_tableH = round(cbind(c(AIC(m5), AIC(m7), AIC(m4), AIC(m1), AIC(m3)),
                                  c(delta1, delta2, delta3, delta4, delta5), 
                                  c(me1, me2, me3, me4, me5), 
                                  c(weight1, weight2, weight3, weight4, weight5)), 3)

eq1 = paste(m5$call)[2]
eq2 = paste(m7$call)[2]
eq3 = paste(m4$call)[2]
eq4 = paste(m1$call)[2]
eq5 = paste(m3$call)[2]

equations = c(eq1, eq2, eq3, eq4, eq5)
equations = rename_regformula(equations, "logHeight", "Site", "Elevation_c", "TSeq_c")

summary_tableH = cbind(equations, summary_tableH)
colnames(summary_tableH) = c("Equation", "AIC","dAIC", "Likelihood","Weight")
```

```{r}
summary_tableH
```
```{r}
# Best fit model summary
cf <- round(coef(m5), 2) 
eqH <- paste0("logHeight = ", cf[1],
             ifelse(sign(cf[2])==1, " + ", " - "), abs(cf[2]), " Site ",
             ifelse(sign(cf[3])==1, " + ", " - "), abs(cf[3]), " TSegIQR_c")

commentH = "Site and TSegIQR_c"
rowH = cbind(eqH, commentH)
colnames(rowH) = c("Best Fit Model", "Significant Vars (p < 0.05)")
rowH
```


```{r}
t.test(ocos$Height~ocos$Site)
t.test(ocos$logHeight~ocos$Site)
```


### Bajada Ocos

**Testing Covariates**

Only significant covariates are listed below:

```{r echo=FALSE}
m1 = lm(Height_c ~ TSegIQR_c, data=ocos_baj)
m2 = lm(logHeight ~ TSegIQR_c, data=ocos_baj)

tidy_regression(m1, output_col) 
tidy_regression(m2, output_col)
```

**Model Comparisons**

**Elevation_c, TSeg_c**

```{r}
data<-data.frame(
  R=ocos_baj$logHeight,
  A=ocos_baj$Elevation_c,
  B=ocos_baj$TSegIQR_c) 

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian lm 2-FF.R")
```

```{r}
anova(m0, m2, test="Chisq") # Adding B does improve fit
anova(m0, m1, test="Chisq") # Adding A does not improve fit
anova(m2, m3, test="Chisq") # Adding A does not improve fit
anova(m3, m4, test="Chisq") # Adding A*B does not improve fit
```

```{r}
tidy_regression(m2, output_col)
tidy_regression(m3, output_col)
tidy_regression(m4, output_col)
tidy_regression(m1, output_col)
tidy_regression(m0, output_col)
```

**Plot diagnostics**

All passed. Residuals are independent.

```{r}
plot_diagnostic(m2, ocos_baj)
plot_diagnostic(m3, ocos_baj)
plot_diagnostic(m4, ocos_baj)
plot_diagnostic(m1, ocos_baj)
```

**Calculate likelihood and weights**

```{r}
# Calculate relative likelihood of each model which is exp( -0.5 * deltaAIC)

delta1 = 0 # abs(AIC(m2) - AIC(m2))
delta2 = abs(AIC(m2) - AIC(m3))
delta3 = abs(AIC(m2) - AIC(m4))
delta4 = abs(AIC(m2) - AIC(m1))
delta5 = abs(AIC(m2) - AIC(m0))


me1 <- exp( -0.5 * delta1)
me2 <- exp( -0.5 * delta2)
me3 <- exp( -0.5 * delta3)
me4 <- exp( -0.5 * delta4)
me5 <- exp( -0.5 * delta5)

# Calculate the Akaike weight for a model which is this value divided by the sum of these values across all models.

sum <- sum(c(me1, me2, me3, me4, me5))

weight1 = me1 / (sum) 
weight2 = me2 / (sum)
weight3 = me3 / (sum)
weight4 = me4 / (sum)
weight5 = me5 / (sum)
```

```{r}
summary_tableH.B = round(cbind(c(AIC(m2), AIC(m3), AIC(m4), AIC(m1), AIC(m0)),
                            c(delta1, delta2, delta3, delta4, delta5), 
                            c(me1, me2, me3, me4, me5), 
                            c(weight1, weight2, weight3, weight4, weight5)), 3)

eq1 = paste(m2$call)[2]
eq2 = paste(m3$call)[2]
eq3 = paste(m4$call)[2]
eq4 = paste(m1$call)[2]
eq5 = paste(m0$call)[2]

equations = c(eq1, eq2, eq3, eq4, eq5)
equations = rename_regformula(equations, "logHeight", "Elevation_c", "TSeq_c")

summary_tableH.B = cbind(equations, summary_tableH.B)
colnames(summary_tableH.B) = c("Equation", "AIC","dAIC", "Likelihood","Weight")
```

```{r}
summary_tableH.B
```

Only factor that largely influences ocotillo height is terminal segment length.

```{r}
# Best fit model summary
cf <- round(coef(m2), 2) 
eqH.B <- paste0("logHeight = ", cf[1],
             ifelse(sign(cf[2])==1, " + ", " - "), abs(cf[2]), " TSeqIQR_c ")

commentH.B = "TSeq_c"
rowH.B = cbind(eqH.B, commentH.B)
colnames(rowH.B) = c("Best Fit Model", "Significant Vars (p < 0.05)")
rowH.B
```

## Circumference

### All Ocos 

**Testing Covariates**

Only significant covariates are listed below:

```{r echo=FALSE}
m1 = lm(Circ_c ~ NBranch_c, data=ocos)
m2 = lm(logCirc ~ NBranch_c, data=ocos)
m3 = lm(logCirc ~ X1m_Num4, data=ocos) # Although significant, this is not a meaningful number in this context.

tidy_regression(m1, output_col) 
tidy_regression(m2, output_col) 
tidy_regression(m3, output_col) 
```

**Plot diagnostics**

All pass. Residuals are independent.

```{r}
plot_diagnostic(m1, ocos)
plot_diagnostic(m2, ocos)
plot_diagnostic(m3, ocos)
```
**Model Comparisons**

Null model was the top model for all other combinations of covariates except this combination:

**NBranches, site**

```{r}
data<-data.frame(
  R=ocos$logCirc,
  A=ocos$NBranch_c,
  B=ocos$Site)

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian lm 2-FF.R")
```

```{r}
anova(m1, m3, test='Chisq') # Adding B does not improve fit
anova(m0, m1, test='Chisq') # Adding A does improve fit
anova(m0, m2, test='Chisq') # Adding B does not improve fit
```

```{r}
tidy_regression(m1, output_col) 
tidy_regression(m0, output_col) 
tidy_regression(m2, output_col)
```

**Plot diagnostics**

All passed. Residuals are independent.

```{r}
plot_diagnostic(m1, ocos)
plot_diagnostic(m0, ocos)
plot_diagnostic(m2, ocos)
```

**Calculate likelihood and weights**

```{r}
# Calculate relative likelihood of each model which is exp( -0.5 * deltaAIC)

delta1 = 0 # abs(AIC(m1) - AIC(m1))
delta2 = abs(AIC(m1) - AIC(m0))
delta3 = abs(AIC(m1) - AIC(m2))

me1 <- exp( -0.5 * delta1)
me2 <- exp( -0.5 * delta2)
me3 <- exp( -0.5 * delta3)

# Calculate the Akaike weight for a model which is this value divided by the sum of these values across all models.

sum <- sum(c(me1, me2, me3))

weight1 = me1 / (sum) 
weight2 = me2 / (sum)
weight3 = me3 / (sum)
```

```{r}
summary_tableC = round(cbind(c(AIC(m1), AIC(m0), AIC(m2)),
                            c(delta1, delta2, delta3), 
                            c(me1, me2, me3), 
                            c(weight1, weight2, weight3)),3)

eq1 = paste(m1$call)[2]
eq2 = paste(m0$call)[2]
eq3 = paste(m2$call)[2]

equations = c(eq1, eq2, eq3)
equations = rename_regformula(equations, "logCirc", "NBranch_c", "Site")

summary_tableC = cbind(equations, summary_tableC)
colnames(summary_tableC) = c("Equation", "AIC","dAIC", "Likelihood","Weight")
```

```{r}
summary_tableC
```


```{r}
# Best fit model summary
cf <- round(coef(m1), 2) 
eqC <- paste0("logCirc = ", cf[1],
             ifelse(sign(cf[2])==1, " + ", " - "), abs(cf[2]), " NBranch_c ")

commentC = "NBranch_c"
rowC = cbind(eqC, commentC)
colnames(rowC) = c("Best Fit Model", "Significant Vars (p < 0.05)")
rowC
```

### Bajada Ocotillos

**Testing Covariates**

Only significant covariates are listed below:

```{r echo=FALSE}
m1 = lm(Circ_c ~ NBranch_c, data=ocos_baj)
m2 = lm(logCirc ~ NBranch_c, data=ocos_baj)
m3 = lm(logCirc ~ X1m_Num4, data=ocos_baj) # AB: although significant is this a meaningful number? 

tidy_regression(m1, output_col) 
tidy_regression(m2, output_col) 
tidy_regression(m3, output_col) 
```

**Plot diagnostics**

All pass. Residuals are independent.

```{r}
plot_diagnostic(m1, ocos_baj)
plot_diagnostic(m2, ocos_baj)
plot_diagnostic(m3, ocos_baj)
```

**Model Comparisons**

Null model was the top model for all other combinations of covariates except this combination:

**NBranches, site, X1m_Num4**

```{r}
data<-data.frame(
  R=ocos_baj$logCirc,
  A=ocos_baj$NBranch_c,
  B=ocos_baj$X1m_Num4) 

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian lm 2-FF.R")
```

```{r}
anova(m2, m3, test="Chisq") # Adding A does improve fit
anova(m1, m3, test="Chisq") # Adding B does improve fit
```

**Plot diagnostics**

All passed. Residuals are independent.

```{r}
plot_diagnostic(m1, ocos_baj)
plot_diagnostic(m2, ocos_baj)
plot_diagnostic(m3, ocos_baj) # but did this pass?
plot_diagnostic(m4, ocos_baj) # but did this pass?
```

Conclusion: Same top model as the all ocotillos model comparisons. Which is lm(formula = logCirc ~ NBranch_c, data = data)

```{r}
# Calculate relative likelihood of each model which is exp( -0.5 * deltaAIC)

delta1 = 0 # abs(AIC(m1) - AIC(m1))
delta2 = abs(AIC(m1) - AIC(m0))

me1 <- exp( -0.5 * delta1)
me2 <- exp( -0.5 * delta2)

# Calculate the Akaike weight for a model which is this value divided by the sum of these values across all models.

sum <- sum(c(me1, me2, me3))

weight1 = me1 / (sum) 
weight2 = me2 / (sum)
```
  
```{r}
summary_tableC.B = round(cbind(c(AIC(m1), AIC(m0)),
                            c(delta1, delta2), c(me1, me2), 
                            c(weight1, weight2)),3)

eq1 = paste(m1$call)[2]
eq2 = paste(m0$call)[2]

equations = c(eq1, eq2)
equations = rename_regformula(equations, "logCirc", "NBranch_c", "X1m_Num4")

summary_tableC.B = cbind(equations, summary_tableC.B)
colnames(summary_tableC.B) = c("Equation", "AIC","dAIC", "Likelihood","Weight")
```

```{r}
summary_tableC.B
```


```{r}
# Best fit model summary
cf <- round(coef(m1), 2) 
eqC.B <- paste0("logCirc = ", cf[1],
             ifelse(sign(cf[2])==1, " + ", " - "), abs(cf[2]), " NBranch_c ")

commentC.B = "NBranch_c"
rowC.B = cbind(eqC.B, commentC.B)
colnames(rowC.B) = c("Best Fit Model", "Significant Vars (p < 0.05)")
rowC.B
```

## Number of Branches

### All Ocos

**Testing Covariates** 

Only significant covariates are listed below:

```{r}
m1 = lm(NBranch_c ~ Circ_c, data=ocos)
m2 = lm(logNB ~ Circ_c, data=ocos)
m3 = lm(NBranch_c ~ Arroyo_c, data=ocos)
m4 = lm(logNB ~ Arroyo_c, data=ocos)
m5 = lm(logNB ~ Inter_Dis_c, data=ocos)
m6 = lm(NBranch_c ~ Inter_plant_b, data=ocos)
m7 = lm(logNB ~ Inter_plant_b, data=ocos)

tidy_regression(m1, output_col) 
tidy_regression(m2, output_col)
tidy_regression(m3, output_col) 
tidy_regression(m4, output_col) 
tidy_regression(m5, output_col) # marginally significant
tidy_regression(m6, output_col) 
tidy_regression(m7, output_col)  
```


**Plot Diagnostics**

All except NBranch ~ Circ passed. However, when the response variable is log transformed, the models pass - residuals are independent.

```{r}
plot_diagnostic(m1, ocos) # did not pass 
plot_diagnostic(m2, ocos) 
plot_diagnostic(m3, ocos) 
plot_diagnostic(m4, ocos)
plot_diagnostic(m5, ocos)
#plot_diagnostic(m6, ocos) # categorical covariate
#plot_diagnostic(m7, ocos) # categorical covariate
```

**Model Comparisons**

Null model was the top model for all other combinations of covariates except this combination:

**Circ_c, Inter_plant_b, Inter_Dis_c, Arroyo_Dis**

```{r}
data<-data.frame(
  R=ocos$logNB,
  A=ocos$Circ_c,
  B=ocos$Inter_plant_b, 
  C=ocos$Inter_Dis_c,
  D=ocos$Arroyo_c) 

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian lm 4-FF.R")
```

```{r}
anova(m13, m32, test='Chisq') # Adding C*D does not improve fit
anova(m13, m15, test='Chisq') # Adding B does not improve fit
anova(m7, m13, test='Chisq') # Adding C marginally improves fit
anova(m10, m13, test='Chisq') # Adding A improves fit
anova(m6, m13, test='Chisq') # Adding D improves fit
```

```{r}
tidy_regression(m13, output_col) # A + C + D
tidy_regression(m10, output_col) # C + D | Add A
tidy_regression(m7, output_col) # A + D | Add C
tidy_regression(m5, output_col)
tidy_regression(m6, output_col) # A + C | Add D
```

**Plot diagnostics**

All passed. Residuals are independent.

```{r}
plot_diagnostic(m13, ocos)
plot_diagnostic(m10, ocos)
plot_diagnostic(m7, ocos)
plot_diagnostic(m5, ocos)
plot_diagnostic(m6, ocos)
```

**Calculate likelihood and weights**

```{r}
# Calculate relative likelihood of each model which is exp( -0.5 * deltaAIC)

delta1 = 0 # abs(AIC(m13) - AIC(m13))
delta2 = abs(AIC(m13) - AIC(m7))
delta3 = abs(AIC(m13) - AIC(m6))
delta4 = abs(AIC(m13) - AIC(m5))

me1 <- exp( -0.5 * delta1)
me2 <- exp( -0.5 * delta2)
me3 <- exp( -0.5 * delta3)
me4 <- exp( -0.5 * delta4)

# Calculate the Akaike weight for a model which is this value divided by the sum of these values across all models.

sum <- sum(c(me1, me2, me3, me4))

weight1 = me1 / (sum) 
weight2 = me2 / (sum)
weight3 = me3 / (sum)
weight4 = me4 / (sum)
```

```{r}
summary_tableB = round(cbind(c(AIC(m13), AIC(m7), AIC(m6), AIC(m5)),
                            c(delta1, delta2, delta3, delta4), 
                            c(me1, me2, me3, me4), 
                            c(weight1, weight2, weight3, weight4)),3)

eq1 = paste(m13$call)[2]
eq2 = paste(m7$call)[2]
eq3 = paste(m6$call)[2]
eq4 = paste(m5$call)[2]

equations = c(eq1, eq2, eq3, eq4)
equations = rename_regformula(equations, "logNB", "Circ_c", "Inter_plant_b", "Inter_Dis_c", "Arroyo_c")

summary_tableB = cbind(equations, summary_tableB)
colnames(summary_tableB) = c("Equation", "AIC","dAIC", "Likelihood","Weight")
```

```{r}
summary_tableB 
```


```{r}
# Best fit model summary
cf <- round(coef(m13), 3) 
eqNB <- paste0("logNB = ", cf[1],
             ifelse(sign(cf[2])==1, " + ", " - "), abs(cf[2]), " Circ_c ",
             ifelse(sign(cf[3])==1, " + ", " - "), abs(cf[3]), " Inter_Dis_c",
             ifelse(sign(cf[4])==1, " + ", " - "), abs(cf[4]), " Arroyo_c")

commentNB = "Circ_c and Arroyo_c; Inter_Dis_c is marginal (p = 0.79)"
rowNB = cbind(eqNB, commentNB)
colnames(rowNB) = c("Best Fit Model", "Significant Vars (p < 0.05)")
rowNB
```

### Bajada Ocos

**Testing Covariates** 

Only significant covariates are listed below:

```{r}
m1 = lm(NBranch_c ~ Circ_c, data=ocos_baj)
m2 = lm(logNB ~ Circ_c, data=ocos_baj)
m3 = lm(NBranch_c ~ Arroyo_c, data=ocos_baj)
m4 = lm(logNB ~ Arroyo_c, data=ocos_baj)
m5 = lm(logNB ~ Inter_Dis_c, data=ocos_baj)
m7 = lm(NBranch_c ~ Elevation_c, data=ocos_baj)
m7 = lm(logNB ~ Elevation_c, data=ocos_baj)
m8 = lm(NBranch_c ~ Arroyo_c, data=ocos_baj)
m9 = lm(logNB ~ Arroyo_c, data=ocos_baj)
m10 = lm(NBranch_c ~ X1m_CID_type1, data=ocos_baj)
m11 = lm(logNB ~ X1m_CID_type1, data=ocos_baj)
m12 = lm(NBranch_c ~ Intra_Dis_c, data=ocos_baj)
m13 = lm(logNB ~ Intra_Dis_c, data=ocos_baj)

tidy_regression(m1, output_col) 
tidy_regression(m2, output_col)
tidy_regression(m3, output_col) 
tidy_regression(m4, output_col) 
tidy_regression(m5, output_col) 
tidy_regression(m6, output_col)  
tidy_regression(m7, output_col) 
tidy_regression(m8, output_col) 
tidy_regression(m9, output_col) 
tidy_regression(m10, output_col) 
tidy_regression(m11, output_col) 
tidy_regression(m12, output_col) 
tidy_regression(m13, output_col) 
```

**Plot Diagnostics**

When the response variable is log transformed, the models passed except for m13 which was lm(logNB ~ NN_c, data=ocos_baj).

```{r}
ocos_baj_s = ocos_baj[complete.cases(ocos_baj$X1m_CID_type1),]
plot_diagnostic(m1, ocos_baj) 
plot_diagnostic(m2, ocos_baj) 
plot_diagnostic(m3, ocos_baj) 
plot_diagnostic(m4, ocos_baj)
plot_diagnostic(m5, ocos_baj)
#plot_diagnostic(m6, ocos_baj) # categorical covariate
plot_diagnostic(m7, ocos_baj) 
plot_diagnostic(m8, ocos_baj) 
plot_diagnostic(m9, ocos_baj) 
#plot_diagnostic(m10, ocos_baj_s) # categorical covariate
#plot_diagnostic(m11, ocos_baj_s) # categorical covariate
plot_diagnostic(m12, ocos_baj) 
plot_diagnostic(m13, ocos_baj) # did not pass
```
**Model Comparisons**

**Elevation_c, Circ_c, Intra_Dis_c**

```{r}
data<-data.frame(
  R=ocos_baj$logNB,
  A=ocos_baj$Elevation_c,
  B=ocos_baj$Circ_c, 
  C=ocos_baj$Intra_Dis_c) 

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian lm 3-FF.R")
```

```{r}
anova(m15, m17, test="Chisq") # Adding A*C does not improve fit
anova(m13, m15, test="Chisq") # Adding A*B marginally improves fit
anova(m10, m13, test="Chisq") # Adding A marginally improves fit
anova(m6, m10, test="Chisq") # Adding B*C improves fit
anova(m6, m7, test="Chisq") # Adding A does not improve fit
anova(m3, m6, test="Chisq") # Adding B does improve fit
anova(m2, m6, test="Chisq") # Adding C does improve fit
```

```{r}
tidy_regression(m10, output_col)
tidy_regression(m13, output_col) 
tidy_regression(m6, output_col) 
tidy_regression(m7, output_col) 
tidy_regression(m2, output_col) 
tidy_regression(m3, output_col) 
```

**Arroyo_c, Circ_c, Intra_Dis_c**

```{r}
data<-data.frame(
  R=ocos_baj$logNB,
  A=ocos_baj$Arroyo_c,
  B=ocos_baj$Circ_c, 
  C=ocos_baj$Intra_Dis_c) 

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian lm 3-FF.R")
```

```{r}
anova(m15, m17, test="Chisq") # Adding A*C does not improve fit
anova(m13, m15, test="Chisq") # Adding A*B marginally improves fit
anova(m10, m13, test="Chisq") # Adding A does not improve fit
anova(m6, m10, test="Chisq") # Adding B*C improves fit
anova(m6, m7, test="Chisq") # Adding A does not improve fit
anova(m3, m6, test="Chisq") # Adding B does improve fit
anova(m2, m6, test="Chisq") # Adding C does improve fit
```

```{r}
tidy_regression(m10, output_col)
tidy_regression(m13, output_col) 
tidy_regression(m6, output_col) 
tidy_regression(m7, output_col) 
tidy_regression(m2, output_col) 
tidy_regression(m3, output_col) 
```

**Inter_plant_b, Circ_c, Intra_Dis_c**

```{r}
data<-data.frame(
  R=ocos_baj$logNB,
  A=ocos_baj$Inter_plant_b,
  B=ocos_baj$Circ_c, 
  C=ocos_baj$Intra_Dis_c) 

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian lm 3-FF.R")
```

```{r}
anova(m15, m17, test="Chisq") # Adding A*C does not improve fit
anova(m13, m15, test="Chisq") # Adding A*B does improve fit
anova(m10, m13, test="Chisq") # Adding A marginally improves fit
anova(m6, m10, test="Chisq") # Adding B*C does improve fit
```

```{r}
tidy_regression(m15, output_col) 
tidy_regression(m17, output_col) 
tidy_regression(m13, output_col) 
tidy_regression(m10, output_col)
tidy_regression(m6, output_col) 
```

**Comparing best fits**

```{r}
m0 = lm(logNB ~ Circ_c * Intra_Dis_c + Inter_plant_b + Arroyo_c, ocos_baj)
m1 = lm(logNB ~ Circ_c * Intra_Dis_c + Inter_plant_b + Elevation_c, ocos_baj)
m2 = lm(logNB ~ Circ_c * Intra_Dis_c + Inter_plant_b, ocos_baj)
m3 =lm(logNB ~ Circ_c * Intra_Dis_c + Elevation_c, ocos_baj)
m4 = lm(logNB ~ Circ_c * Intra_Dis_c + Arroyo_c, ocos_baj)
m5 = lm(logNB ~ Circ_c * Intra_Dis_c, ocos_baj)
m6 = lm(logNB ~ Circ_c + Intra_Dis_c, ocos_baj)
m7 = lm(logNB ~ Circ_c, ocos_baj)
m8 = lm(logNB ~ Intra_Dis_c, ocos_baj)
```

```{r}
anova(m2, m1, test="Chisq") # Adding Elevation_c does not improve fit
anova(m2, m0, test="Chisq") # Adding Arroyo_c does not improve fit
```

```{r}
tidy_regression(m2, output_col) # top model
```

**Plot diagnostics**

All passed except m8. Residuals are independent.

```{r}
plot_diagnostic(m2, ocos_baj)
plot_diagnostic(m1, ocos_baj)
plot_diagnostic(m5, ocos_baj)
plot_diagnostic(m0, ocos_baj)
plot_diagnostic(m3, ocos_baj)
plot_diagnostic(m4, ocos_baj)
plot_diagnostic(m6, ocos_baj)
plot_diagnostic(m7, ocos_baj)
plot_diagnostic(m8, ocos_baj) # does not pass confidently
```

**Calculate likelihood and weights**

```{r}
# Calculate relative likelihood of each model which is exp( -0.5 * deltaAIC)

delta1 = 0 # abs(AIC(m13) - AIC(m13))
delta2 = abs(AIC(m2) - AIC(m1))
delta3 = abs(AIC(m2) - AIC(m5))
delta4 = abs(AIC(m2) - AIC(m0))
delta5 = abs(AIC(m2) - AIC(m3))

me1 <- exp( -0.5 * delta1)
me2 <- exp( -0.5 * delta2)
me3 <- exp( -0.5 * delta3)
me4 <- exp( -0.5 * delta4)
me5 <- exp( -0.5 * delta5)

# Calculate the Akaike weight for a model which is this value divided by the sum of these values across all models.

sum <- sum(c(me1, me2, me3, me4, me5))

weight1 = me1 / (sum) 
weight2 = me2 / (sum)
weight3 = me3 / (sum)
weight4 = me4 / (sum)
weight5 = me5 / (sum)
```

```{r}
summary_tableB.B = round(cbind(c(AIC(m2), AIC(m1), AIC(m5), AIC(m0), AIC(m3)),
                            c(delta1, delta2, delta3, delta4, delta5), 
                            c(me1, me2, me3, me4, me5), 
                            c(weight1, weight2, weight3, weight4, weight5)),3)

eq1 = paste(m2$call)[2]
eq2 = paste(m1$call)[2]
eq3 = paste(m5$call)[2]
eq4 = paste(m0$call)[2]
eq5 = paste(m3$call)[2]

equations = c(eq1, eq2, eq3, eq4, eq5)

summary_tableB.B = cbind(equations, summary_tableB.B)
colnames(summary_tableB.B) = c("Equation", "AIC","dAIC", "Likelihood","Weight")
```

```{r}
summary_tableB.B
```


```{r}
# Best fit model summary
cf <- round(coef(m2), 2) 
eqNB.B <- paste0("logNB = ", cf[1],
             ifelse(sign(cf[2])==1, " + ", " - "), abs(cf[2]), " Circ_c ",
             ifelse(sign(cf[3])==1, " + ", " - "), abs(cf[3]), " Intra_Dis_c",
             ifelse(sign(cf[4])==1, " + ", " - "), abs(cf[4]), " Inter_plant_b",
             ifelse(sign(cf[5])==1, " + ", " - "), abs(cf[5]), " Circ_c*Intra_Dis_c")

commentNB.B = "Circ_c and Circ_c*Intra_Dis_c"
rowNB.B = cbind(eqNB.B, commentNB.B)
colnames(rowNB.B) = c("Best Fit Model", "Significant Vars (p < 0.05)")
rowNB.B
```

```{r}
t.test(ocos$logNB~ocos$Inter_Plant_Group)
```

# Mixed Effect Modeling of Ocotillo Terminal Segments

Analyses below are mixed effect models of Fouquieria splendens terminal segment length for all individuals located on both a bajada and a plain in Organ Pipe National Monument, Arizona. Terminal segment length denotes the length of each of the last five, topmost branches of the ten branches measured per ocotillo. The best fit model is the model with the lowest AIC value. The rest of the models were ordered by their ascending AIC values. No subsets of the data were used in these analyses.

## Re-Reading and Cleaning the Data

```{r}
source("src/cleaning_data.R")
source("src/regression_output.R")
output_col = FALSE

all_data = clean_data("data/General_Oco_Data.csv", "data/Terminal_5Segs.csv")

seg_baj = all_data[[1]][[1]]
segs = all_data[[1]][[2]]

segs <- segs[!is.na(segs$Length_c),]
```

## Functions

```{r}
### Diagnostic Plot Function

plot_diagnostic = function(m, d) {
  # m:  model generated by lm() or glm()
  # d:  data as a data.frame object
  
  d['fitted_values'] = fitted(m)
  d['fitted_residuals'] = residuals(m)
  
  plot = ggplot(data = d) + geom_point(aes(x = fitted_values, y = fitted_residuals)) +
  geom_hline(yintercept = 0) +
  labs(title = "Plot of Fitted Residuals against Fitted Values",
       x = "Fitted Values", y = "Fitted Residuals") 
  
  return(plot)
}
```

```{r}
## Renaming Regression Formulas Function

rename_regformulaME = function (eqs, response_var, predA, predB, predC, randX, randY) {
  # eqs:              list of linear regression formals as characters
  # response_var:     response variable as a character
  # predA:            predictor variable A as a character
  # predB:            predictor variable B as a character
  # predC:            predictor variable C as a character
  # randX:            random variable X as a character
  # randY:            random variable Y as a character
  for (i in 1:length(eqs)) {
  if (grepl(" A", eqs[i], fixed = TRUE)) {
    eqs[i] =  gsub("A", predA, eqs[i])
  }
  if (grepl(" B", eqs[i], fixed = TRUE)) {
    eqs[i] =  gsub("B", predB, eqs[i])
  }
  if (grepl(" C", eqs[i], fixed = TRUE)) {
    eqs[i] =  gsub("C", predC, eqs[i])
  }
  if (grepl("X", eqs[i], fixed = TRUE)) {
    eqs[i] =  gsub("X", randX, eqs[i])
  }
    if (grepl("Y", eqs[i], fixed = TRUE)) {
    eqs[i] =  gsub("Y", randY, eqs[i])
  }
  if (grepl( "R", eqs[i], fixed = TRUE)) {
    eqs[i] = gsub("R", response_var, eqs[i])
  }
  }
  return(eqs)
}
```

**Testing Covariates**

Dropped branch_num as a random factor to better help the model fit the data.

Only significant covariates are listed below (most are marginal in this case):

```{r echo=FALSE}
m1 = lmer(Length_c ~ Height_c + (1|Tree) + (1|seg_num), data=segs)
m2 = lmer(Length_c ~ T1m_NumCacti + (1|Tree) + (1|seg_num), data=segs) 
m3 = lmer(Length_c ~ R1m_shrub2cactus + (1|Tree) + (1|seg_num), data=segs)

# lmer(Length_c ~ X1m_type2 + (1|Tree) + (1|seg_num), data=segs) also significant but it does not answer our question on how the number and types of small competitors affect terminal segment length.

tidy_regression(m1, output_col) 
tidy_regression(m2, output_col)
tidy_regression(m3, output_col)
```

**Plot diagnostics**

All passed. Residuals are independent.

```{r}
temp_segs = segs[complete.cases(segs$R1m_shrub2cactus),]

plot_diagnostic(m1, segs)
plot_diagnostic(m2, segs)
plot_diagnostic(m3, temp_segs)
#plot_diagnostic(m4, temp_segs)
```

**Model Comparisons**

**Height, T1m_NumCacti, 1m_shrub2cactus**

```{r}
data<-data.frame(
  R=segs$Length_c,
  A=segs$Height_c, 
  B=segs$T1m_NumCacti, 
  C=segs$R1m_shrub2cactus,
  X=segs$Tree,
  Y=segs$seg_num)

# Fit the models to the same number of observations by filtering out NA values:
data <- data[complete.cases(data),]

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian glmer 2-RF + 3-FF REMLF.R")
```

```{r}
anova(m4, m7, test="Chisq") # Adding C does not improve fits
anova(m1, m4, test="Chisq") # Adding B does not improve fit
anova(m1, m5, test="Chisq") # Adding C does not improve fit
anova(m4, m8, test="Chisq") # Adding A*B does not improve fit
anova(m0, m1, test="Chisq") # Adding A does improve fit
```

```{r}
tidy_regression(m4, output_col)
tidy_regression(m1, output_col)
tidy_regression(m5, output_col)
tidy_regression(m0, output_col)
```

**Plot diagnostics**

All passed. Residuals are independent.

```{r}
plot_diagnostic(m4, data)
plot_diagnostic(m1, data)
plot_diagnostic(m5, data)
plot_diagnostic(m0, data)
```

**Calculate likelihood and weights**

```{r}
# Calculate relative likelihood of each model which is exp( -0.5 * deltaAIC)

delta1 = 0 # abs(AIC(m4) - AIC(m4))
delta2 = abs(AIC(m4) - AIC(m1))
delta3 = abs(AIC(m4) - AIC(m5))
delta4 = abs(AIC(m4) - AIC(m0))

me1 <- exp( -0.5 * delta1)
me2 <- exp( -0.5 * delta2)
me3 <- exp( -0.5 * delta3)
me4 <- exp( -0.5 * delta4)

# Calculate the Akaike weight for a model which is this value divided by the sum of these values across all models.

sum <- sum(c(me1, me2, me3, me4))

weight1 = me1 / (sum) 
weight2 = me2 / (sum)
weight3 = me3 / (sum)
weight4 = me4 / (sum)
```


```{r}
summary_tableS = round(cbind(c(AIC(m4), AIC(m1), AIC(m5), AIC(m0)),
                            c(delta1, delta2, delta3, delta4), 
                            c(me1, me2, me3, me4), 
                            c(weight1, weight2, weight3, weight4)),3)

eq1 = c(paste(formula(m4))[2], paste(formula(m4))[1], paste(formula(m4))[3]) %>%
 paste(collapse=' ')
eq2 = c(paste(formula(m1))[2], paste(formula(m1))[1], paste(formula(m1))[3]) %>%
 paste(collapse=' ')
eq3 = c(paste(formula(m5))[2], paste(formula(m5))[1], paste(formula(m5))[3]) %>%
 paste(collapse=' ')
eq4 = c(paste(formula(m0))[2], paste(formula(m0))[1], paste(formula(m0))[3]) %>%
 paste(collapse=' ')

equations = c(eq1, eq2, eq3, eq4)
equations = rename_regformulaME(equations, "Length_c", "Height_c", "T1m_NumCacti", "1m_shrub2cactus", "Tree", "seg_num")

summary_tableS = cbind(equations, summary_tableS)
colnames(summary_tableS) = c("Equation", "AIC","dAIC", "Likelihood","Weight")
```

```{r}
summary_tableS
```


```{r}
# Best fit model summary
cf <- round(summary(m4)$coefficient, 2)
eqS <- paste0("Length_c = ", cf[1],
             ifelse(sign(cf[2])==1, " + ", " - "), abs(cf[2]), " Height_c",
             ifelse(sign(cf[3])==1, " + ", " - "), abs(cf[3]), " T1m_NumCacti", 
             " + (1 | Tree) + (1 | seg_num)")

commentS = "Height_c"
rowS = cbind(eqS, commentS)
colnames(rowS) = c("Best Fit Model", "Significant Vars (p < 0.05)")
rowS
```


## Bajada Ocos

```{r}
seg_baj = seg_baj[!is.na(seg_baj$Length),]
```

**Testing Covariates**

Dropped branch_num as a random factor to better help the model fit the data.

Only significant covariates are listed below (most are marginal in this case):

```{r echo=FALSE}
m1 = lmer(Length_c ~ Height_c + (1|Tree) + (1|seg_num), data=seg_baj)
m2 = lmer(Length_c ~ T1m_NumCacti + (1|Tree) + (1|seg_num), data=seg_baj) 
m3 = lmer(Length_c ~ R1m_shrub2cactus + (1|Tree) + (1|seg_num), data=seg_baj)

tidy_regression(m1, output_col) 
tidy_regression(m2, output_col)
tidy_regression(m3, output_col)
```
**Plot diagnostics**

All passed. Residuals are independent.

```{r}
temp_seg_baj = seg_baj[complete.cases(seg_baj$R1m_shrub2cactus),]

plot_diagnostic(m1, seg_baj)
plot_diagnostic(m2, seg_baj)
plot_diagnostic(m3, temp_seg_baj)
```

**Height, T1m_NumCacti, 1m_shrub2cactus**

```{r}
data<-data.frame(
  R=seg_baj$Length_c,
  A=seg_baj$Height_c, 
  B=seg_baj$T1m_NumCacti, 
  C=seg_baj$R1m_shrub2cactus,
  X=seg_baj$Tree,
  Y=seg_baj$seg_num)

# Fit the models to the same number of observations by filtering out NA values:
data <- data[complete.cases(data),]

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian glmer 2-RF + 3-FF REMLF.R")
```
```{r}
anova(m1, m4, test="Chisq") # Adding B does improve fit
anova(m1, m5, test="Chisq") # Adding C does not improve fit
```

```{r}
tidy_regression(m4, output_col)
tidy_regression(m1, output_col)
tidy_regression(m5, output_col)
tidy_regression(m0, output_col)
```

**Plot diagnostics**

All passed. Residuals are independent.

```{r}
plot_diagnostic(m4, data)
plot_diagnostic(m1, data)
plot_diagnostic(m5, data)
plot_diagnostic(m0, data)
```

**Calculate likelihood and weights**

```{r}
# Calculate relative likelihood of each model which is exp( -0.5 * deltaAIC)

delta1 = 0 # abs(AIC(m4) - AIC(m4))
delta2 = abs(AIC(m4) - AIC(m1))
delta3 = abs(AIC(m4) - AIC(m5))
delta4 = abs(AIC(m4) - AIC(m0))

me1 <- exp( -0.5 * delta1)
me2 <- exp( -0.5 * delta2)
me3 <- exp( -0.5 * delta3)
me4 <- exp( -0.5 * delta4)

# Calculate the Akaike weight for a model which is this value divided by the sum of these values across all models.

sum <- sum(c(me1, me2, me3, me4))

weight1 = me1 / (sum) 
weight2 = me2 / (sum)
weight3 = me3 / (sum)
weight4 = me4 / (sum)
```

```{r}
summary_tableS.B = round(cbind(c(AIC(m4), AIC(m5), AIC(m1), AIC(m0)),
                            c(delta1, delta2, delta3, delta4), 
                            c(me1, me2, me3, me4), 
                            c(weight1, weight2, weight3, weight4)),3)

eq1 = c(paste(formula(m4))[2], paste(formula(m4))[1], paste(formula(m4))[3]) %>%
 paste(collapse=' ')
eq2 = c(paste(formula(m5))[2], paste(formula(m5))[1], paste(formula(m5))[3]) %>%
 paste(collapse=' ')
eq3 = c(paste(formula(m1))[2], paste(formula(m1))[1], paste(formula(m1))[3]) %>%
 paste(collapse=' ')
eq4 = c(paste(formula(m0))[2], paste(formula(m0))[1], paste(formula(m0))[3]) %>%
 paste(collapse=' ')

equations = c(eq1, eq2, eq3, eq4)
equations = rename_regformulaME(equations, "Length_c", "Height_c", "T1m_NumCacti", "1m_shrub2cactus", "Tree", "seg_num")

summary_tableS.B = cbind(equations, summary_tableS.B)
colnames(summary_tableS) = c("Equation", "AIC","dAIC", "Likelihood","Weight")
```

```{r}
summary_tableS.B
```


```{r}
# Best fit model summary
cf <- round(summary(m4)$coefficient, 2)
eqS.B <- paste0("Length_c = ", cf[1],
             ifelse(sign(cf[2])==1, " + ", " - "), abs(cf[2]), " Height_c",
             ifelse(sign(cf[3])==1, " + ", " - "), abs(cf[3]), " T1m_NumCacti", 
             " + (1 | Tree) + (1 | seg_num)")

commentS.B = "Height_c and T1m_NumCacti"  
rowS.B = cbind(eqS.B, commentS.B)
colnames(rowS.B) = c("Best Fit Model", "Significant Vars (p < 0.05)")
rowS.B
```


## Summary Tables to CSV

```{r}
dataset = c("all", "bajada", "all", "bajada", "all", "bajada", "all", "bajada")
best_fit_table = rbind(rowH, rowH.B,
                       rowS, rowS.B,
                       rowC, rowC.B,
                       rowNB, rowNB.B)
best_fit_table = cbind(dataset, best_fit_table)

best_fit_table
write.csv(best_fit_table,'best_fits_morph.csv', row.names=FALSE)
```


```{r}
setS = c("all", "all", "all", "all",
             "bajada", "bajada", "bajada", "bajada")
setH = c("all", "all", "all", "all", "all",
             "bajada", "bajada", "bajada", "bajada", "bajada")
setC = c("all", "all", "all",
             "bajada", "bajada")
setB = c("all", "all", "all", "all",
             "bajada", "bajada", "bajada", "bajada", "bajada")

datasets = c(setH, setS, setC, setB)
```

```{r}
full_table = rbind(summary_tableH, summary_tableH.B,
                   summary_tableS, summary_tableS.B,
                   summary_tableC, summary_tableC.B,
                   summary_tableB, summary_tableB.B)

full_table = cbind(datasets, full_table)
colnames(full_table) = c("Ocotillos", "Equation", "AIC","dAIC", "Likelihood","Weight")
full_table

write.csv(full_table,'morph_regressions.csv', row.names=FALSE)
```

## Kruskal-Wallis 

**Re-Reading and Cleaning the Data**

```{r}
source("src/cleaning_data.R")
all_data = clean_data("data/General_Oco_Data.csv", "data/Terminal_5Segs.csv")

seg_baj = all_data[[1]][[1]]
segs = all_data[[1]][[2]]

segs <- segs[!is.na(segs$Length_c),]
```


**Histogram**

```{r fig.width=8, fig.height=3.5}
par(mfrow=c(1,2))
hist(segs$Length, main="", xlab= "Terminal Segment Length (cm)", cex.lab=1.6, cex.axis=1.4, xlim=c(0,80), ylim=c(0,500), col="white")
title("(a)", adj = 0.05, line = 0, cex.main=1.5)
```


```{r}
kruskal.test(Length ~ Inter_Plant_Group, data = segs)
table <- pairwise.wilcox.test(segs$Length, segs$COMP_id,
                 p.adjust.method = "BH")

model1 <- kruskal.test(Length ~ Inter_Plant_Group, data = segs)
model2 <- kruskal.test(Length ~ Inter_ID, data = segs)

#table$p.value

summary(aov(segs$Length ~ segs$Inter_ID))
competitor <- segs$Inter_ID
model = aov(segs$Length ~ competitor)
tukey <- TukeyHSD(x=model, conf.level=0.95)
par(mar = c(4, 14, 4, 1) , cex.axis=.7)
plot(tukey , las=1 , col="brown")
```
**By Competitor Species**

```{r fig.height=4, fig.width=5.3}
data = data.frame(Segment_Length=segs$Length, 
                  Inter_Species=segs$Inter_ID, 
                  Inter_Plant_Group=segs$Inter_Plant_Group)
pwc <- segs %>% 
  dunn_test(Length ~ Inter_ID, p.adjust.method = "bonferroni")
pwc <- pwc %>% add_xy_position(x = "group")

res.kruskal <- segs %>% kruskal_test(Length ~ Inter_ID)
res.kruskal

ggplot(segs, aes(Inter_ID, Length)) + 
  geom_violin() +
    labs(
    subtitle = get_test_label(res.kruskal, detailed = TRUE), 
    caption = get_pwc_label(pwc)
    ) + 
  theme(legend.position="none") +
  xlab("Major Nearest Shrub Competitor Species") +
  ylab("Terminal Segment Length (cm)") +
  geom_boxplot(width=0.1) +
  stat_pvalue_manual(pwc, hide.ns = TRUE)
```

**By Competitor Type**

```{r fig.width=4, fig.height=3.5}
pwc <- segs %>% 
  dunn_test(Length ~ Inter_Plant_Group, p.adjust.method = "bonferroni")
pwc <- pwc %>% add_xy_position(x = "Inter_Plant_Group")

res.kruskal <- segs %>% kruskal_test(Length ~ Inter_Plant_Group)
res.kruskal

ggplot(segs, aes(Inter_Plant_Group, Length)) + 
  geom_violin() +
    labs(
    subtitle = get_test_label(res.kruskal, detailed = TRUE)
    ) + 
  theme(legend.position="none") +
  xlab("Nearest Major Interspecific Group") +
  ylab("Terminal Segment Length (cm)") +
  geom_boxplot(width=0.1) +
  theme(text = element_text(size=20)) +
  stat_pvalue_manual(pwc, hide.ns = TRUE)
```

**By Interspecies Group Comparing Bajada v.s. Plain**

```{r fig.width=4, fig.height=3.5}

#### Bajada by Interspecies Group

pwc <- seg_baj %>% 
  dunn_test(Length ~ Inter_Plant_Group, p.adjust.method = "bonferroni")
pwc <- pwc %>% add_xy_position(x = "Inter_Plant_Group")

res.kruskal <- seg_baj %>% kruskal_test(Length ~ Inter_Plant_Group)
res.kruskal

ggplot(seg_baj, aes(Inter_Plant_Group,Length)) + 
  geom_violin() +
    labs(
    subtitle = get_test_label(res.kruskal, detailed = TRUE)
    ) + 
  theme(legend.position="none") +
  xlab("Nearest Major Interspecies Group") +
  ylab("Terminal Segment Length (cm)") +
  geom_boxplot(width=0.1) +
  theme(text = element_text(size=20)) +
  stat_pvalue_manual(pwc, hide.ns = TRUE)

#### Plain by Interspecies Group

seg_plain = segs[segs$Site == 1,]

pwc <- seg_plain %>% 
  dunn_test(Length ~ Inter_Plant_Group, p.adjust.method = "bonferroni")
pwc <- pwc %>% add_xy_position(x = "Inter_Plant_Group")

res.kruskal <- seg_plain %>% kruskal_test(Length ~ Inter_Plant_Group)
res.kruskal

ggplot(seg_plain, aes(Inter_Plant_Group,Length)) + 
  geom_violin() +
    labs(
    subtitle = get_test_label(res.kruskal, detailed = TRUE)
    ) + 
  theme(legend.position="none") +
  xlab("Nearest Major Interspecies Group") +
  ylab("Terminal Segment Length (cm)") +
  geom_boxplot(width=0.1) +
  theme(text = element_text(size=20)) +
  stat_pvalue_manual(pwc, hide.ns = TRUE)
```


**By Site **

```{r}
segs$Site_b = segs$Site
segs$Site[segs$Site_b == 0] <- "bajada"
segs$Site[segs$Site_b == 1] <- "plain"
```


```{r fig.width=4, fig.height=3.5}
segs$Site_b = as.factor(segs$Site)

pwc <- segs %>% 
  dunn_test(Length ~ Site, p.adjust.method = "bonferroni")
pwc <- pwc %>% add_xy_position(x = "Site")

res.kruskal <- segs %>% kruskal_test(Length ~ Site)
res.kruskal

ggplot(segs, aes(Site, Length)) + 
  geom_violin() +
    labs(
    subtitle = get_test_label(res.kruskal, detailed = TRUE)
    ) + 
  theme(legend.position="none") +
  xlab("Site Type") +
  ylab("Terminal Segment Length (cm)") +
  geom_boxplot(width=0.1) +
  theme(text = element_text(size=20)) +
  stat_pvalue_manual(pwc, hide.ns = TRUE)
```


## Principal Component Analysis (PCA)

**Reading the Data**

```{r}
growthForm <- read.csv("data/growthForm.csv", row.names = 1)
```

**Investigating Variables**

```{r}
GFpca <- PCA(growthForm, scale.unit = TRUE, graph = TRUE, ncp = 10)

eig.val <- get_eigenvalue(GFpca)
eig.val
fviz_eig(GFpca, addlabels = TRUE, ylim = c(0, 50))
```

An eigenvalue > 1 indicates that PCs account for more variance than accounted by one of the original variables in standardized data. This is commonly used as a cutoff point for which PCs are retained. This holds true only when the data are standardized, which we have done.

You can also limit the number of component to that number that accounts for a certain fraction of the total variance. (No formal cut-off)

```{r}
variables <- get_pca_var(GFpca)
corrplot(variables$cos2, is.corr=FALSE)

corrplot(variables$contrib, is.corr=FALSE)

GFdim <- dimdesc(GFpca, axes = c(1,2), proba = 0.05)
GFdim$Dim.1
GFdim$Dim.2
```

**Plot by Group**

```{r}
Groupings <- read_csv("data/Groupings.csv")
growthForm$Inter_Plant_Group <- Groupings$Inter_Plant_Group
growthForm$Site <- Groupings$Site

fviz_pca_ind(GFpca,
             geom.ind = "text",
             col.ind = growthForm$Inter_Plant_Group,
             legend.title = "Interspecific Group",
             addEllipses = TRUE
)

fviz_pca_ind(GFpca,
             geom.ind = "text",
             col.ind = growthForm$Site,
             legend.title = "Site Type",
              addEllipses = TRUE
)
```


