---
title: "Ocotillo Neighbor Summary Statistics"
author: "Anastasia Bernat"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
rm(list=ls())

library(ggplot2)
library(olsrr)

dir = "/Users/anastasiabernat/Desktop/git_repositories/ocotillo-research/"
setwd(dir)

knitr::opts_chunk$set(echo = TRUE)
```

# Multivariate Modeling of Ocotillo Inter- and Intra-Species Distance

Analyses below are multivariate models of distance to the nearest neighboring Fouquieria splendens and nearest interspecies neighbor for ocotillos located on both a bajada and a plain in Organ Pipe National Monument, Arizona. All models were grouped by their response variable and ordered by their ascending AIC values. Dataset "ocos" indicates all individuals measured on the bajada and plain while "ocos_baj" indicates only the individuals measured on the bajada. Interspecies neighbor group is split between two types – shrub and cactus – where cactus = 0 and shrub = 1. 

## Reading and Cleaning the Data

```{r}
source("src/cleaning_data.R")
source("src/regression_output.R")
output_col = FALSE

all_data = clean_data("data/General_Oco_Data.csv", "data/End_Segments_5segs.csv")

ocos = all_data[[2]]
ocos_baj = ocos[ocos$Plant[1:20],]
```

## Functions

```{r}
### Diagnostic Plot Function

plot_diagnostic = function(m, d) {
  
  d['fitted_values'] = m$fitted.values
  d['fitted_residuals'] = m$residuals
  
  plot = ggplot(data = d) + geom_point(aes(x = fitted_values, y = fitted_residuals)) +
  geom_hline(yintercept = 0) +
  labs(title = "Plot of Fitted Residuals against Fitted Values",
       x = "Fitted Values", y = "Fitted Residuals") 
  
  print(ols_test_normality(m))
  
  return(plot)
}
```

```{r}
### Renaming Regression Formulas Function

rename_regformula = function (eqs, response_var, predA, predB, predC, predD) {
  # eqs:              list of linear regression formals as characters
  # response_var:     response variable as a character
  # predA:            predictor variable A as a character
  # predB:            predictor variable B as a character
  # predC:            predictor variable C as a character
  # predD:            predictor variable D as a character
  for (i in 1:length(eqs)) {
  if (grepl(" A", eqs[i], fixed = TRUE)) {
    eqs[i] =  gsub("A", predA, eqs[i])
  }
  if (grepl(" B", eqs[i], fixed = TRUE)) {
    eqs[i] =  gsub("B", predB, eqs[i])
  }
  if (grepl(" C", eqs[i], fixed = TRUE)) {
    eqs[i] =  gsub("C", predC, eqs[i])
  }
  if (grepl(" D", eqs[i], fixed = TRUE)) {
    eqs[i] =  gsub("D", predD, eqs[i])
  }
  if (grepl( "R", eqs[i], fixed = TRUE)) {
    eqs[i] = gsub("R", response_var, eqs[i])
  }
  }
  return(eqs)
}
```

## Nearest Neighbor 

### All Ocos

**Testing Covariates**

Only significant covariates are listed below:

```{r echo=FALSE}
m1 = lm(logNN ~ Elevation_c, data=ocos)
m2 = lm(NN_c ~ Elevation_c, data=ocos)
m3 = lm(logNN ~ site, data=ocos)
m4 = lm(NN_c ~ site, data=ocos)
m5 = lm(logNN ~ Arroyo_c, data=ocos)
m6 = lm(NN_c ~ Arroyo_c, data=ocos)

tidy_regression(m1, output_col) 
tidy_regression(m2, output_col)
tidy_regression(m3, output_col)
tidy_regression(m4, output_col) 
tidy_regression(m5, output_col)
tidy_regression(m6, output_col)
```

When the response variable is log transformed, residuals are independent.

```{r}
plot_diagnostic(m1, ocos)
plot_diagnostic(m2, ocos)
#plot_diagnostic(m3, ocos) # categorical covariate
#plot_diagnostic(m4, ocos) # categorical covariate
plot_diagnostic(m5, ocos)
plot_diagnostic(m5, ocos)
```

**Model Comparisons**

**site, Elevation_c, Circ**

```{r}
data<-data.frame(
  R=ocos$logNN,
  A=ocos$site,
  B=ocos$Elevation_c, 
  C=ocos$Circ_c) 

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian lm 3-FF.R")
```

```{r}
cor(ocos$site, ocos$Elevation_c)
```

Elevation and site were highly correlated (R = -0.85), which leads to multicollinearity. In turn, interactions between site and elevation were removed. This eliminated m11.

```{r}
anova(m2, m6, test='Chisq') # Adding C marginally improves fit
anova(m6, m7, test='Chisq') # Adding A does not improve fit
anova(m0, m2, test='Chisq') # Adding B improves fit
```

**Height, Elevation_c, Circ**

```{r}
data<-data.frame(
  R=ocos$logNN,
  A=ocos$Height_c,
  B=ocos$Elevation_c, 
  C=ocos$Circ_c) 

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian lm 3-FF.R")
```

```{r}
anova(m16, m17, test='Chisq') # Adding A*C does not improve fit
anova(m12, m16, test='Chisq') # Adding B*C does improve fit
```

**Height, Arroyo_c, Circ**

```{r}
data<-data.frame(
  R=ocos$logNN,
  A=ocos$Height_c,
  B=ocos$Arroyo_c, 
  C=ocos$Circ_c) 

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian lm 3-FF.R")
```

```{r}
anova(m12, m14, test='Chisq') # Adding A*B does not improve fit
anova(m9, m12, test='Chisq') # Adding B marginally improves fit 
anova(m5, m9, test='Chisq') # Adding A*C improves fit
```

**Comparing best fits**

```{r}
m0 = lm(logNN ~ Height_c * Circ_c, data=ocos)
m1 = lm(logNN ~ Height_c + Circ_c, data=ocos)
m2 = lm(logNN ~ Height_c * Circ_c + Arroyo_c , data=ocos)
m3 = lm(logNN ~ Height_c * Circ_c  + Elevation_c * Circ_c , data = ocos)
m4 = lm(logNN ~ Height_c * Circ_c  + Elevation_c, data = ocos)
```

**Plot diagnostics**

All passed. Residuals are independent.

```{r}
plot_diagnostic(m1, ocos)
plot_diagnostic(m0, ocos)
plot_diagnostic(m2, ocos)
plot_diagnostic(m4, ocos)
plot_diagnostic(m3, ocos)
```

**Calculate likelihood and weights**

```{r}
# Calculate relative likelihood of each model which is exp( -0.5 * deltaAIC)

delta1 = 0 # abs(AIC(m5) - AIC(m5))
delta2 = abs(AIC(m3) - AIC(m4))
delta3 = abs(AIC(m3) - AIC(m2))
delta4 = abs(AIC(m3) - AIC(m0))
delta5 = abs(AIC(m3) - AIC(m1))

me1 <- exp( -0.5 * delta1)
me2 <- exp( -0.5 * delta2)
me3 <- exp( -0.5 * delta3)
me4 <- exp( -0.5 * delta4)
me5 <- exp( -0.5 * delta5)

# Calculate the Akaike weight for a model which is this value divided by the sum of these values across all models.

sum <- sum(c(me1, me2, me3, me4, me5))

weight1 = me1 / (sum) 
weight2 = me2 / (sum)
weight3 = me3 / (sum)
weight4 = me4 / (sum)
weight5 = me5 / (sum)
```

```{r}
summary_tableN = round(cbind(c(AIC(m3), AIC(m4), AIC(m2), AIC(m0), AIC(m1)),
                                  c(delta1, delta2, delta3, delta4, delta5), 
                                  c(me1, me2, me3, me4, me5), 
                                  c(weight1, weight2, weight3, weight4, weight5)), 3)

eq1 = paste(m3$call)[2]
eq2 = paste(m4$call)[2]
eq3 = paste(m2$call)[2]
eq4 = paste(m0$call)[2]
eq5 = paste(m1$call)[2]

equations = c(eq1, eq2, eq3, eq4, eq5)

summary_tableN = cbind(equations, summary_tableN)
colnames(summary_tableN) = c("Equation", "AIC","dAIC", "Likelihood","Weight")
```

```{r}
summary_tableN
```

```{r}
# Best fit model summary
cf <- round(coef(m3), 2) 
eqNN <- paste0("logNN = ", cf[1],
             ifelse(sign(cf[2])==1, " + ", " - "), abs(cf[2]), " Height",
             ifelse(sign(cf[3])==1, " + ", " - "), abs(cf[3]), " Circ_c",
             ifelse(sign(cf[4])==1, " + ", " - "), abs(cf[4]), " Elevation_c",
             ifelse(sign(cf[5])==1, " + ", " - "), abs(cf[5]), " Height_c*Circ_c",
             ifelse(sign(cf[6])==1, " + ", " - "), abs(cf[6]), " Circ_c*Elevation_c")

commentNN = "Circ_c and Height_c*Circ_c | Elevation_c and Circ_c*Elevation_c are marginally sig (p < 0.10)"
rowNN = cbind(eqNN, commentNN)
colnames(rowNN) = c("Best Fit Model", "Significant Vars (p < 0.05)")
rowNN
```


### Bajada Ocos

**Testing Covariates**

Only significant covariates are listed below:

```{r echo=FALSE}
m1 = lm(logNN ~ Elevation_c, data=ocos_baj)
m2 = lm(NN_c ~ Elevation_c, data=ocos_baj)
m3 = lm(logNN ~ NBranch_c, data=ocos_baj)
m4 = lm(NN_c ~ NBranch_c, data=ocos_baj)
m5 = lm(logNN ~ COMP_plant_b, data=ocos_baj)
m6 = lm(NN_c ~ COMP_plant_b, data=ocos_baj)
m7 = lm(logNN ~ COMP_dis_c, data=ocos_baj)
m8 = lm(logNN ~ Circ_c, data=ocos_baj)
m9 = lm(NN_c ~ Circ_c, data=ocos_baj)
m10 = lm(logNN ~ Arroyo_c, data=ocos_baj)
m11 = lm(NN_c ~ Arroyo_c, data=ocos_baj)

tidy_regression(m1, output_col) 
tidy_regression(m2, output_col)
tidy_regression(m3, output_col)
tidy_regression(m4, output_col)
tidy_regression(m5, output_col)
tidy_regression(m6, output_col)
tidy_regression(m7, output_col) 
tidy_regression(m8, output_col) 
tidy_regression(m9, output_col)
tidy_regression(m10, output_col)
tidy_regression(m11, output_col)
```

**Plot diagnostics**

All log-transformed response variables passed. Residuals were independent.

```{r}
plot_diagnostic(m1, ocos_baj)
plot_diagnostic(m2, ocos_baj)
plot_diagnostic(m3, ocos_baj)
plot_diagnostic(m4, ocos_baj)
plot_diagnostic(m5, ocos_baj)
plot_diagnostic(m6, ocos_baj)
plot_diagnostic(m7, ocos_baj)
plot_diagnostic(m8, ocos_baj)
plot_diagnostic(m9, ocos_baj)
plot_diagnostic(m10, ocos_baj)
plot_diagnostic(m11, ocos_baj)
```

**Model Comparisons**

**Elevation_c, NBranch_c, COMP_dis_c**

```{r}
data<-data.frame(
  R=ocos_baj$logNN,
  A=ocos_baj$Elevation_c,
  B=ocos_baj$NBranch_c,
  C=ocos_baj$COMP_dis_c) 

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian lm 3-FF.R")
```

```{r}
anova(m13, m15, test="Chisq") # Adding A*B does not improve fit
anova(m10, m13, test="Chisq") # Adding A does improve fit
```

```{r}
tidy_regression(m13, output_col)
tidy_regression(m10, output_col)
```

**Arroyo_c, NBranch_c, COMP_dis_c**

```{r}
data<-data.frame(
  R=ocos_baj$logNN,
  A=ocos_baj$Arroyo_c,
  B=ocos_baj$NBranch_c,
  C=ocos_baj$COMP_dis_c) 

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian lm 3-FF.R")
```

```{r}
anova(m13, m16, test='Chisq') # Adding A*C does not improve fit
anova(m10, m13, test='Chisq') # Adding A does improve fit
```

```{r}
tidy_regression(m13, output_col)
tidy_regression(m10, output_col)
```

**Comparing best fits**

```{r}
m1 = lm(logNN ~  NBranch_c * COMP_dis_c + Elevation_c, data = ocos_baj)
m2 = lm(logNN ~ NBranch_c * COMP_dis_c, data = ocos_baj)
m3 = lm(logNN ~  NBranch_c* COMP_dis_c + Arroyo_c, data = ocos_baj)
m4 = lm(logNN ~  NBranch_c + COMP_dis_c, data = ocos_baj)
```

**Plot diagnostics**

All passed. Residuals are independent.

```{r}
plot_diagnostic(m1, ocos_baj)
plot_diagnostic(m2, ocos_baj)
plot_diagnostic(m3, ocos_baj)
plot_diagnostic(m4, ocos_baj)
```

**Calculate likelihood and weights**

```{r}
# Calculate relative likelihood of each model which is exp( -0.5 * deltaAIC)

delta1 = 0 # abs(AIC(m2) - AIC(m2))
delta2 = abs(AIC(m1) - AIC(m3))
delta3 = abs(AIC(m1) - AIC(m2))
delta4 = abs(AIC(m1) - AIC(m4))

me1 <- exp( -0.5 * delta1)
me2 <- exp( -0.5 * delta2)
me3 <- exp( -0.5 * delta3)
me4 <- exp( -0.5 * delta4)

# Calculate the Akaike weight for a model which is this value divided by the sum of these values across all models.

sum <- sum(c(me1, me2, me3, me4))

weight1 = me1 / (sum) 
weight2 = me2 / (sum)
weight3 = me3 / (sum)
weight4 = me4 / (sum)
```

```{r}
summary_tableN.B = round(cbind(c(AIC(m1), AIC(m3), AIC(m2), AIC(m4)),
                                  c(delta1, delta2, delta3, delta4), 
                                  c(me1, me2, me3, me4), 
                                  c(weight1, weight2, weight3, weight4)), 3)

eq1 = paste(m1$call)[2]
eq2 = paste(m3$call)[2]
eq3 = paste(m2$call)[2]
eq4 = paste(m4$call)[2]

equations = c(eq1, eq2, eq3, eq4)

summary_tableN.B = cbind(equations, summary_tableN.B)
colnames(summary_tableN.B) = c("Equation", "AIC","dAIC", "Likelihood","Weight")
```

```{r}
summary_tableN.B
```


```{r}
# Best fit model summary
cf <- round(coef(m1), 2) 
eqNN.B <- paste0("logNN = ", cf[1],
             ifelse(sign(cf[2])==1, " + ", " - "), abs(cf[2]), " NBranch_c",
             ifelse(sign(cf[3])==1, " + ", " - "), abs(cf[3]), " COMP_dis_c",
             ifelse(sign(cf[4])==1, " + ", " - "), abs(cf[4]), " Elevation_c",
             ifelse(sign(cf[5])==1, " + ", " - "), abs(cf[5]), " NBranch_c*COMP_dis_c")

commentNN.B = "Elevation_c and NBranch_c*COMP_dis_c"
rowNN.B = cbind(eqNN.B, commentNN.B)
colnames(rowNN.B) = c("Best Fit Model", "Significant Vars (p < 0.05)")
rowNN.B
```


## Nearest (Major) Interspecies Neighbor

### All Ocos 

**Testing Covariates**

Only significant covariates are listed below:

```{r echo=FALSE}
m1 = lm(logCOMP ~ Arroyo_c, data=ocos)
m2 = lm(COMP_dis_c ~ Arroyo_c, data=ocos)
m3 = lm(COMP_dis_c ~ Elevation_c, data=ocos)

tidy_regression(m1, output_col) 
tidy_regression(m2, output_col) 
tidy_regression(m3, output_col) 
```

**Plot diagnostics**

All pass. Residuals are independent.

```{r}
plot_diagnostic(m1, ocos)
plot_diagnostic(m2, ocos)
plot_diagnostic(m3, ocos)
```

**Model Comparisons**

Null model was the top model for all other combinations of covariates except this combination:

**Arroyo_c, Elevation**

```{r}
data<-data.frame(
  R=ocos$logCOMP,
  A=ocos$Arroyo_c,
  B=ocos$Elevation_c) 

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian lm 2-FF.R")
```

```{r}
anova(m0, m1, test="Chisq") # Adding A does improve fit
anova(m0, m2, test="Chisq") # Adding B does not improve fit
```

```{r}
tidy_regression(m0, output_col) 
tidy_regression(m2, output_col)
tidy_regression(m1, output_col) 
```

**Plot diagnostics**

All passed. Residuals are independent.

```{r}
plot_diagnostic(m0, ocos)
plot_diagnostic(m2, ocos)
plot_diagnostic(m1, ocos)
```

**Calculate likelihood and weights**

```{r}
# Calculate relative likelihood of each model which is exp( -0.5 * deltaAIC)

delta1 = 0 # abs(AIC(m1) - AIC(m1))
delta2 = abs(AIC(m1) - AIC(m2))
delta3 = abs(AIC(m1) - AIC(m0))

me1 <- exp( -0.5 * delta1)
me2 <- exp( -0.5 * delta2)
me3 <- exp( -0.5 * delta3)

# Calculate the Akaike weight for a model which is this value divided by the sum of these values across all models.

sum <- sum(c(me1, me2, me3))

weight1 = me1 / (sum) 
weight2 = me2 / (sum)
weight3 = me3 / (sum)
```
  
```{r}
summary_tableC = round(cbind(c(AIC(m1), AIC(m2), AIC(m0)),
                                  c(delta1, delta2, delta3), 
                                  c(me1, me2, me3), 
                                  c(weight1, weight2, weight3)), 3)

eq1 = paste(m1$call)[2]
eq2 = paste(m2$call)[2]
eq3 = paste(m0$call)[2]

equations = c(eq1, eq2, eq3)
equations = rename_regformula(equations, "logCOMP", "Arroyo_c", "Elevation_c")

summary_tableC = cbind(equations, summary_tableC)
colnames(summary_tableC) = c("Equation", "AIC","dAIC", "Likelihood","Weight")
```

```{r}
summary_tableC
```

```{r}
# Best fit model summary
cf <- round(coef(m1), 3) 
eqC <- paste0("logCOMP = ", cf[1],
             ifelse(sign(cf[2])==1, " + ", " - "), abs(cf[2]), " Arroyo_c")

commentC = "Arroyo is marginally significant (p=0.054)"
rowC = cbind(eqC, commentC)
colnames(rowC) = c("Best Fit Model", "Significant Vars (p < 0.05)")
rowC
```

### Bajada Ocotillos

**Testing Covariates**

Only significant covariates are listed below:

```{r echo=FALSE}
m1 = lm(logCOMP ~ Arroyo_c, data=ocos_baj)
m2 = lm(COMP_dis_c ~ Arroyo_c, data=ocos_baj)
m3 = lm(logCOMP ~ Elevation_c, data=ocos_baj)
m4 = lm(COMP_dis_c ~ Elevation_c, data=ocos_baj)

tidy_regression(m1, output_col) 
tidy_regression(m2, output_col) 
tidy_regression(m3, output_col) 
tidy_regression(m4, output_col) 
```

**Plot diagnostics**

All pass. Residuals are independent.

```{r}
plot_diagnostic(m1, ocos_baj)
plot_diagnostic(m2, ocos_baj)
plot_diagnostic(m3, ocos_baj)
plot_diagnostic(m4, ocos_baj)
```

**Model Comparisons**

Null model was the top model for all other combinations of covariates except this combination:

**Elevation, Watershed**

```{r}
data<-data.frame(
  R=ocos_baj$logCOMP,
  A=ocos_baj$Arroyo_c,
  B=ocos_baj$Elevation_c) 

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian lm 2-FF.R")
```

```{r}
anova(m0, m2, test="Chisq") # Adding B does improve fit
anova(m0, m1, test="Chisq") # Adding A does improve fit
anova(m1, m3, test="Chisq") # Adding B does not improve fit
```

```{r}
tidy_regression(m0, output_col) 
tidy_regression(m1, output_col) 
tidy_regression(m2, output_col)
```

**Plot diagnostics**

All passed. Residuals are independent.

```{r}
plot_diagnostic(m0, ocos_baj)
plot_diagnostic(m1, ocos_baj)
plot_diagnostic(m2, ocos_baj)
```

**Calculate likelihood and weights**

```{r}
# Calculate relative likelihood of each model which is exp( -0.5 * deltaAIC)

delta1 = 0 # abs(AIC(m1) - AIC(m1))
delta2 = abs(AIC(m1) - AIC(m2))
delta3 = abs(AIC(m1) - AIC(m0))

me1 <- exp( -0.5 * delta1)
me2 <- exp( -0.5 * delta2)
me3 <- exp( -0.5 * delta3)

# Calculate the Akaike weight for a model which is this value divided by the sum of these values across all models.

sum <- sum(c(me1, me2, me3))

weight1 = me1 / (sum) 
weight2 = me2 / (sum)
weight3 = me3 / (sum)
```
  
```{r}
summary_tableC.B = round(cbind(c(AIC(m1), AIC(m2), AIC(m0)),
                                  c(delta1, delta2, delta3), 
                                  c(me1, me2, me3), 
                                  c(weight1, weight2, weight3)), 3)

eq1 = paste(m1$call)[2]
eq2 = paste(m2$call)[2]
eq3 = paste(m0$call)[2]

equations = c(eq1, eq2, eq3)
equations = rename_regformula(equations, "logCOMP", "Arroyo_c", "Elevation_c")

summary_tableC.B = cbind(equations, summary_tableC.B)
colnames(summary_tableC.B) = c("Equation", "AIC","dAIC", "Likelihood","Weight")
```

```{r}
summary_tableC.B
```


```{r}
# Best fit model summary
cf <- round(coef(m1), 3) 
eqC.B <- paste0("logCOMP = ", cf[1],
             ifelse(sign(cf[2])==1, " + ", " - "), abs(cf[2]), " Arroyo_c")

commentC.B = "Arroyo_c"
rowC.B = cbind(eqC.B, commentC.B)
colnames(rowC.B) = c("Best Fit Model", "Significant Vars (p < 0.05)")
rowC.B
```


## Summary Tables to CSV

```{r}
dataset = c("all", "bajada", "all", "bajada")
best_fit_table = rbind(rowNN, rowNN.B,
                       rowC, rowC.B)
best_fit_table = cbind(dataset, best_fit_table)

best_fit_table
write.csv(best_fit_table,'best_fits_neighbor.csv', row.names=FALSE)
```

```{r}
datasets = c("all", "all", "all", "all", "all", 
             "bajada", "bajada", "bajada", "bajada", 
             "all", "all", "all",
             "bajada", "bajada", "bajada")

full_table = rbind(summary_tableN, summary_tableN.B, 
                   summary_tableC, summary_tableC.B)

full_table = cbind(datasets, full_table)
colnames(full_table) = c("Ocotillos", "Equation", "AIC","dAIC", "Likelihood","Weight")

full_table
write.csv(full_table,'neighbor_regressions.csv', row.names=FALSE)
```


