---
title: "PCA Exploration"
author: "Acacia and Anastasia"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())

#load packages
require(FactoMineR)
require(factoextra)
require(corrplot)
require(readr)
library(corrplot)

knitr::opts_chunk$set(echo = TRUE)
```

**Reading and Cleaning the Data**

```{r}
growthForm <- read.csv("data/growthForm.csv", row.names = 1)
growthForm <- read.csv("data/growthForm2.csv", row.names = 1)
Groupings <- read_csv("data/Groupings.csv")
```

```{r}
source("src/cleaning_data.R")
source("src/regression_output.R")
output_col = FALSE

all_data = clean_data("data/General_Oco_Data.csv", "data/Terminal_5Segs.csv")

ocos = all_data[[2]]
ocos_baj = ocos[ocos$Plant[1:20],]

seg_baj = all_data[[1]][[1]]
segs = all_data[[1]][[2]]

segs <- segs[!is.na(segs$Length_c),]
```

**Oultier Visualization**

```{r}
c1 = 1.2 * 1.2
c2 = 1.3 * 1.2
c3 = 1.5 * 1.2
c4 = 1.1 * 1.2
```

```{r}
outlier = ocos[-1,]
x=d[,"TSegIQR_c"]
y=d[,"logHeight"] 

m <- lm(logHeight ~ TSegIQR_c, data=outlier)
x.seq = seq(min(x) - sd(x), max(x) + sd(x), length.out=100)
m <- lm(y ~ x, data=outlier)
prd <- data.frame(x=x.seq) # newdata
err <- predict(m, newdata = prd, se.fit = TRUE)
prd$lci <- err$fit - 1.96 * err$se.fit
prd$fit <- err$fit
prd$uci <- err$fit + 1.96 * err$se.fit
mu_ci <- t(matrix(c(prd$lci,prd$uci), ncol=2))

colfunc <- colorRampPalette(c("red", "orange"))
plot(outlier$logHeight ~ outlier$TSegIQR_c, 
     col=colfunc(2)[as.factor(outlier$Site)], 
     pch = c(17:16)[as.factor(outlier$Site)],
     xlab = "Terminal Segment Length IQR (cm)",
     ylab= "Log(Height) (m)", 
     cex=c2,
     cex.lab=c3,
     cex.axis=c2,
     bty="n") 
legend(7,1,
       legend = c("Bajada", "Plain"),
       col= c(colfunc(2)[1], colfunc(2)[2]),
       pch = c(17:16),
       cex=c4,
       title="Site Type")
abline(m, lty=2)
shade(mu_ci, lim = prd$x)
pval <- paste0("p = ", round(summary(m)$coefficients[8],3))
text(1, 1.5, pval, cex=c1)
title("(b)", adj = 0.05, line = 0, cex.main=c3)

summary(lm(logHeight ~ TSegIQR_c + Site, data=outlier))
```

**Cleveland dotplot**

```{r}
par(mfrow=c(1,2))
x = ocos$Terminal_SegIQR
y = seq(1, length(x),)

plot(x,y, ylab="Order of the data", xlab="Terminal Segment Length IQR")
boxplot(ocos$Terminal_SegIQR)
```

**Revisiting PCA**

```{r}
#for plotting
Round <- function(number){
  x <- round(number, 1)
  if(x%%1 == 0){
    return(paste(as.character(x), ".0", sep = ""))
  }
  else{
    return(x)
  }
}

```

```{r}
#The actual thing
PCA_graphs <- function(dataset){
    GFpca <- PCA(dataset, scale.unit = TRUE, graph = TRUE, ncp = 10)
    
    eig.val <- get_eigenvalue(GFpca)
    var.val <- GFpca$var
    print(eig.val) #will only show in in console
    print(var.val)
    
    scree <- fviz_eig(GFpca, addlabels = TRUE, ylim = c(0, 50))
    print(scree)
    
    labX <- paste("PC1 (", Round(eig.val[1, 2]), "%)", sep = "")
    labY <- paste("PC1 (", Round(eig.val[2, 2]), "%)", sep = "")
    leplot <- fviz_pca_biplot(GFpca, geom.id = c("point"), geom.var = c("arrow", "text"), label = "var", repel = TRUE, col.ind = "gray", col.var = "black")
    ggpubr::ggpar(leplot,title = "Principal Component Analysis of F. splendens Morphology", xlab = labX, ylab = labY,ggtheme = theme_classic(), font.main = 16, font.x = 14, font.y = 14, font.tickslab = 12)
}
```


**PCA and Corrplot Exploration**

```{r}
colnames(growthForm)
```

```{r}
# Everything
d = growthForm
PCA_graphs(d)

colnames(d) = c("Height", "Circ", "NB", "M TSL", "TSL IQR", "BL IQR", "M BL", "Nodes")
D = cor(d)
test <- cor.mtest(d)$p
corrplot.mixed(D,lower.col = "black", number.cex = .7, p.mat=test, sig.level=0.05)
corrplot.mixed(D,lower.col = "black", number.cex = .7)
```

```{r}
# Plant 1 Removed 
d = growthForm[-1,]
PCA_graphs(d)

colnames(d) = c("Height", "Circ", "NB", "M TSL", "TSL IQR", "M BL", "Nodes")
D = cor(d)
test <- cor.mtest(d)$p
corrplot.mixed(D,lower.col = "black", number.cex = .7, p.mat=test, sig.level=0.05)
corrplot.mixed(D,lower.col = "black", number.cex = .7)
```

```{r}
# Plant 1 Removed | Without Branch Length IQR 
d = growthForm[-1,-6]
PCA_graphs(d)

colnames(d) = c("Height", "Circ", "NB", "M TSL", "TSL IQR", "M BL", "Nodes")
D = cor(d)
test <- cor.mtest(d)$p
corrplot.mixed(D,lower.col = "black", number.cex = .7, p.mat=test, sig.level=0.05)
corrplot.mixed(D,lower.col = "black", number.cex = .7)
```
```{r}
# Plant 1 Removed | Without Number Nodes 
d = growthForm[-1,-8]
PCA_graphs(d)

colnames(d) = c("Height", "Circ", "NB", "M TSL", "TSL IQR", "BL IQR", "M BL")
D = cor(d)
test <- cor.mtest(d)$p
corrplot.mixed(D,lower.col = "black", number.cex = .7, p.mat=test, sig.level=0.05)
corrplot.mixed(D,lower.col = "black", number.cex = .7)
```

**Log Transformed**

```{r}
# Everything
d = log(growthForm)
PCA_graphs(d)

colnames(d) = c("Height", "Circ", "NB", "M TSL", "TSL IQR", "BL IQR", "M BL", "Nodes")
D = cor(d)
test <- cor.mtest(d)$p
corrplot.mixed(D,lower.col = "black", number.cex = .7, p.mat=test, sig.level=0.05)
corrplot.mixed(D,lower.col = "black", number.cex = .7)
```

```{r}
# Plant 1 Removed 
d = log(growthForm[-1,])
PCA_graphs(d)

colnames(d) = c("Height", "Circ", "NB", "M TSL", "TSL IQR", "M BL", "Nodes")
D = cor(d)
test <- cor.mtest(d)$p
corrplot.mixed(D,lower.col = "black", number.cex = .7, p.mat=test, sig.level=0.05)
corrplot.mixed(D,lower.col = "black", number.cex = .7)
```

```{r}
# Plant 1 Removed | Without Branch Length IQR 
# AB: This seems the most reasonable looking one...
d = log(growthForm[-1,-6])
PCA_graphs(d)

colnames(d) = c("Height", "Circ", "NB", "M TSL", "TSL IQR", "M BL", "Nodes")
D = cor(d)
test <- cor.mtest(d)$p
corrplot.mixed(D,lower.col = "black", number.cex = .7, p.mat=test, sig.level=0.05)
corrplot.mixed(D,lower.col = "black", number.cex = .7)
```

```{r}
# Plant 1 Removed | Without Number Nodes 
d = log(growthForm[-1,-8])
PCA_graphs(d)

colnames(d) = c("Height", "Circ", "NB", "M TSL", "TSL IQR", "BL IQR", "M BL")
D = cor(d)
test <- cor.mtest(d)$p
corrplot.mixed(D,lower.col = "black", number.cex = .7, p.mat=test, sig.level=0.05)
corrplot.mixed(D,lower.col = "black", number.cex = .7)
```

**Regressions**

```{r}
ocos = all_data[[2]]
```

```{r}
ocos$BL_IQR = growthForm$Branch.Length.IQR
ocos$Median_BL = growthForm$Median.Branch.Length
ocos$NumNodes = growthForm$Number.Nodes
```

### All Ocos

**Testing Covariates**

Only significant covariates are listed below:

```{r}
ocos = ocos[-1,]
```

```{r echo=FALSE}
m1 = lm(logHeight ~ Elevation_c, data=ocos)
m2 = lm(Height_c ~ Site, data=ocos)
m3 = lm(logHeight ~ Site, data=ocos)
m4 = lm(Height_c ~ TSegIQR_c, data=ocos)
m5 = lm(logHeight ~ TSegIQR_c, data=ocos)
m6 = lm(Height ~ BL_IQR, data=ocos)
m7 = lm(logHeight ~ BL_IQR, data=ocos)
m8 = lm(Height ~ Median_BL, data=ocos)
m9 = lm(logHeight ~ Median_BL, data=ocos)
m10 = lm(Height ~ NumNodes, data=ocos)
m11 = lm(logHeight ~ NumNodes, data=ocos)

tidy_regression(m1, output_col) 
tidy_regression(m2, output_col)
tidy_regression(m3, output_col)
tidy_regression(m4, output_col) # not sig when remove outlier
tidy_regression(m5, output_col) # not sig when remove outlier
tidy_regression(m6, output_col) # not sig when remove outlier
tidy_regression(m7, output_col) # not sig when remove outlier
tidy_regression(m8, output_col) 
tidy_regression(m9, output_col)
tidy_regression(m10, output_col) # sig when remove outlier
tidy_regression(m11, output_col) # sign when remove outlier

# Elevation_c, Site, Median_BL, NumNodes no matter if remove outlier
```

**Model Comparisons**

**site, Elevation_c, Median_BL**

```{r}
data<-data.frame(
  R=ocos$logHeight,
  A=ocos$Site,
  B=ocos$Elevation_c, 
  C=ocos$Median_BL) 

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian lm 3-FF.R")
```
```{r}
summary(m5)
```

**Site, NumNodes, TSeg_c**

```{r}
data<-data.frame(
  R=ocos$logHeight,
  A=ocos$Site,
  B=ocos$NumNodes, 
  C=ocos$Median_BL) 

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian lm 3-FF.R")
```
```{r}
anova(m5, m7, test="Chisq") # Adding B does not improve fit
```


```{r fig.width=2.1, fig.height=5.18}

par(mfrow=c(3,1))
###### Plot A

d = ocos
x=d[,"Median_BL"]
y=d[,"logHeight"] 

m <- lm(logHeight ~ Median_BL, data=ocos)
x.seq = seq(min(x) - sd(x), max(x) + sd(x), length.out=100)
m <- lm(y ~ x, data=ocos)
prd <- data.frame(x=x.seq) # newdata
err <- predict(m, newdata = prd, se.fit = TRUE)
prd$lci <- err$fit - 1.96 * err$se.fit
prd$fit <- err$fit
prd$uci <- err$fit + 1.96 * err$se.fit
mu_ci <- t(matrix(c(prd$lci,prd$uci), ncol=2))

colfunc <- colorRampPalette(c("red", "orange"))
plot(ocos$logHeight ~ ocos$Median_BL, 
     col=colfunc(2)[as.factor(ocos$Site)], 
     pch = c(17:16)[as.factor(ocos$Site)],
     xlab = "Median Branch Length (cm)",
     ylab= "Log(Height) (m)", 
     cex=c2,
     cex.lab=c3,
     cex.axis=c2,
     bty="n") 
legend(300,0.95,
       legend = c("Bajada", "Plain"),
       col= c(colfunc(2)[1], colfunc(2)[2]),
       pch = c(17:16),
       cex=c4,
       title="Site Type")
abline(m, lty=2)
shade(mu_ci, lim = prd$x)
pval <- paste0("p = ", round(summary(m)$coefficients[8],5), "*")
text(240, 1.25, pval, cex=c1)
title("(a)", adj = 0.05, line = 0, cex.main=c3)

###### Plot B

d = ocos
x=d[,"BL_IQR"]
y=d[,"logHeight"] 

m <- lm(logHeight ~ BL_IQR, data=ocos)
x.seq = seq(min(x) - sd(x), max(x) + sd(x), length.out=100)
m <- lm(y ~ x, data=ocos)
prd <- data.frame(x=x.seq) # newdata
err <- predict(m, newdata = prd, se.fit = TRUE)
prd$lci <- err$fit - 1.96 * err$se.fit
prd$fit <- err$fit
prd$uci <- err$fit + 1.96 * err$se.fit
mu_ci <- t(matrix(c(prd$lci,prd$uci), ncol=2))

colfunc <- colorRampPalette(c("red", "orange"))
plot(ocos$logHeight ~ ocos$BL_IQR, 
     col=colfunc(2)[as.factor(ocos$Site)], 
     pch = c(17:16)[as.factor(ocos$Site)],
     xlab = "Branch Length Interquartile Range (cm)",
     ylab= "Log(Height) (m)", 
     cex=c2,
     cex.lab=c3,
     cex.axis=c2,
     bty="n") 
legend(110,0.95,
       legend = c("Bajada", "Plain"),
       col= c(colfunc(2)[1], colfunc(2)[2]),
       pch = c(17:16),
       cex=c4,
       title="Site Type")
abline(m, lty=2)
shade(mu_ci, lim = prd$x)
pval <- paste0("p = ", round(summary(m)$coefficients[8],2), "*")
text(100, 1.4, pval, cex=c1)
title("(b)", adj = 0.05, line = 0, cex.main=c3)

###### Plot C

ocos_baj = ocos[2:20,]
d = ocos_baj
x=d[,"BL_IQR"]
y=d[,"logHeight"] 

m <- lm(logHeight ~ BL_IQR, data=ocos_baj)
x.seq = seq(min(x) - sd(x), max(x) + sd(x), length.out=100)
m <- lm(y ~ x, data=ocos_baj)
prd <- data.frame(x=x.seq) # newdata
err <- predict(m, newdata = prd, se.fit = TRUE)
prd$lci <- err$fit - 1.96 * err$se.fit
prd$fit <- err$fit
prd$uci <- err$fit + 1.96 * err$se.fit
mu_ci <- t(matrix(c(prd$lci,prd$uci), ncol=2))

colfunc <- colorRampPalette(c("red", "orange"))
plot(ocos_baj$logHeight ~ ocos_baj$BL_IQR, 
     col=colfunc(2)[as.factor(ocos_baj$Site)], 
     pch = c(17:16)[as.factor(ocos_baj$Site)],
     xlab = "Median Branch Length (cm)",
     ylab= "Log(Height) (m)", 
     cex=c2,
     cex.lab=c3,
     cex.axis=c2,
     bty="n") 
legend(110,0.85,
       legend = c("Bajada", "Plain"),
       col= c(colfunc(2)[1], colfunc(2)[2]),
       pch = c(17:16),
       cex=c4,
       title="Site Type")
abline(m, lty=2)
shade(mu_ci, lim = prd$x)
pval <- paste0("p = ", round(summary(m)$coefficients[8],2))
text(100, 1, pval, cex=c1)
title("(c)", adj = 0.05, line = 0, cex.main=c3)

```

### Bajada Ocos

```{r}
ocos = all_data[[2]]
ocos_baj = ocos[ocos$Plant[1:20],]
```

```{r}
ocos$BL_IQR = growthForm$Branch.Length.IQR
ocos$Median_BL = growthForm$Median.Branch.Length
ocos$NumNodes = growthForm$Number.Nodes
```

**Testing Covariates**

Only significant covariates are listed below:

```{r}
ocos_baj = ocos[1:20,]
ocos_baj = ocos[2:20,]
```

```{r echo=FALSE}
m1 = lm(Height_c ~ TSegIQR_c, data=ocos_baj) # not sig after remove outlier
m2 = lm(logHeight ~ TSegIQR_c, data=ocos_baj) # not sig after remove outlier
m3 = lm(Height ~ Median_BL, data=ocos_baj)
m4 = lm(logHeight ~ Median_BL, data=ocos_baj)
m5 = lm(Height ~ BL_IQR, data=ocos_baj) # not sig after remove outlier
m6 = lm(logHeight ~ BL_IQR, data=ocos_baj) # not sig after remove outlier

tidy_regression(m1, output_col) 
tidy_regression(m2, output_col)
tidy_regression(m3, output_col)
tidy_regression(m4, output_col)
tidy_regression(m5, output_col)
tidy_regression(m6, output_col)
```

**Model Comparisons**

**site, Elevation_c, Median_BL**

```{r}
data<-data.frame(
  R=ocos$logHeight,
  A=ocos$Site,
  B=ocos$Elevation_c, 
  C=ocos$Median_BL) 

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian lm 3-FF.R")
```
```{r}
anova(m5, m7, test="Chisq") # Adding B does not improve fit
```

**site, NumNodes, Median_BL**

```{r}
data<-data.frame(
  R=ocos$logHeight,
  A=ocos$Site,
  B=ocos$NumNodes, 
  C=ocos$Median_BL) 

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian lm 3-FF.R")
```

```{r}
anova(m7, m13, test="Chisq") # adding B*C does not improve fit
anova(m5, m7, test="Chisq") # Adding B does not improve fit | marginal
```
